Draft

### 2019-05-28 draft.tex
Added lemma 3.3 and talked a bit about IP

\usepackage{answers}

\usepackage{setspace}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{tikz-cd}

\usepackage{multicol}
\usepackage{mathrsfs}
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb}
\usepackage{ textcomp }
\usepackage{ wasysym }
\usepackage{bookmark}
\usepackage{accents}
\usepackage[parfill]{parskip}

@@

\newenvironment{bprooftree}
  {\leavevmode\hbox\bgroup}
  {\DisplayProof\egroup}

@@

\maketitle
%Below is an example of the problem environment
\section{Type Theory}
\subsection{Introduction}
Type theory is a formal language and deductive system, that is self sufficient in the sense that it need not be formulated as a collection of axioms on top of some other formal system like First Order Logic, instead its deductive system can be built on top of its own formal language.

@@

\section{LLPO}
\begin{defn} 
The Lesser Limited Principle of Omniscience, states that given binary sequence $s : \mathbb{N} \rightarrow \mathbf{2}$ and the fact that there is at most one occurence of $1$ in the sequence, formally 
$$\mathtt{atMost1one} :\prod_{n_1 : \mathbb{N}} \prod_{n_2 : \mathbb{N}} s(n_1) = 1 \rightarrow s(n_2)= 1 \rightarrow n_1 = n_2$$
we can then have by the LLPO a witness for $\mathtt{p_{odd}} \vee \mathtt{p_{even}}$, where $\mathtt{p_{odd}}$ is the statement that for all odd positions $n$, $s(n) = 0$, formally $\mathtt{p_{odd}} \equiv \prod_{n : \mathbb{N}} \mathtt{odd}(n) \rightarrow s(n) = 0$. 
Similarly for $\mathtt{p_{even}}$. 
\end{defn}
LLPO can be viewed as a weaker form of the Law of Excluded Middle.

@@

	\item $\neg \mathtt{p}$, alongside LEM, implies that there exists odd $n_e : \mathbb{N}$ such that $s(n_e) = 1$. 
	We can now provide $\prod_{n : \mathbb{N}} \mathtt{even}(n) \rightarrow s(n) = 0$. 
	Let $n : \mathbb{N}$. 
	By the definition of $s$, $s(n) = 0 \vee s(n) = 1$.
	We invoke the principle of induction of coproducts and prove LLPO from the disjuncts. 
	\begin{itemize}
		\item $s(n) = 0$, in this case we are done.
		\item $s(n) = 1$. 
		By $\mathtt{atMost1one}$ we have $n = n_e \Rightarrow $ $n$ even and odd which is a contradiction $c : \bot$. 
		Ex Falso,  $\mathtt{efq} : \bot \rightarrow s(n) = 0$. 
		Then $\mathtt{efq}(c) : s(n) = 0$.
	\end{itemize}

>>>>

%\usepackage{answers}

\usepackage{setspace}
%\usepackage{graphicx}
\usepackage{enumitem}
%\usepackage{tikz-cd}
%\usepackage{todonotes}
\usepackage{multicol}
\usepackage{mathrsfs}
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb}
\usepackage{ textcomp }
\usepackage{ wasysym }
\usepackage{bookmark}
\usepackage{accents}
\usepackage[parfill]{parskip}

@@

\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}

%\newenvironment{bprooftree}
 % {\leavevmode\hbox\bgroup}
  %{\DisplayProof\egroup}

@@

\maketitle
%Below is an example of the problem environment
\section{Type Theory}
\subsection{Introduction}
Type theory is a formal language and deductive system, that is self sufficient in the sense that it need not be formulated as a collection of axioms on top of some other formal system like First Order Logic, instead its deductive system can be built on top of its own formal language. 
\todo{longer sentence to check formating }

@@

\section{LLPO}
\begin{defn} 
The Lesser Limited Principle of Omniscience, states that given binary sequence $s : \mathbb{N} \rightarrow \mathbf{2}$ and the fact that there is at most one occurence of $1$ in the sequence, formally 
$$\mathtt{atMost1one} :\prod_{n_1 : \mathbb{N}} \prod_{n_2 : \mathbb{N}} s(n_1) = 1 \rightarrow s(n_2)= 1 \rightarrow n_1 = n_2$$
we can then have by the LLPO a witness for $\mathtt{p_{odd}} \vee \mathtt{p_{even}}$, where $\mathtt{p_{odd}}$ (with $s$ as an implicit argument) is the statement that for all odd positions $n$, $s(n) = 0$, formally $\mathtt{p_{odd}} \equiv \prod_{n : \mathbb{N}} \mathtt{odd}(n) \rightarrow s(n) = 0$. 
Similarly for $\mathtt{p_{even}}$. 
\end{defn}
LLPO can be viewed as a weaker form of the Law of Excluded Middle.

@@

	\item $\neg \mathtt{p_{odd}}$, alongside LEM, implies that there exists odd $n_e : \mathbb{N}$ such that $s(n_e) = 1$. 
	We can now provide $\prod_{n : \mathbb{N}} \mathtt{even}(n) \rightarrow s(n) = 0$. 
	Let $n : \mathbb{N}$. 
	By the definition of $s$, $s(n) = 0 \vee s(n) = 1$.
	We invoke the principle of induction of coproducts and prove LLPO from the disjuncts. 
	\begin{itemize}
		\item $s(n) = 0$, in this case we are done.
		\item $s(n) = 1$. 
		By $\mathtt{atMost1one}$ we have $n = n_e \Rightarrow $ $n$ even and odd which is a contradiction $c : \bot$. 
		Ex Falso,  $\mathtt{efq} : \bot \rightarrow s(n) = 0$. 
		Then $\mathtt{efq}(c) : s(n) = 0$.
	\end{itemize}
\end{itemize}
\end{proof}

\begin{lem}
If we replace the consequent of LLPO with its double negation, let's call it $\neg\neg$LLPO, then we can prove it in Type Theory. 
\end{lem}
\begin{proof}
From LEM$\;\Rightarrow\;$LLPO we have LEM$\;\Rightarrow\neg\neg$LLPO by effectively the same proof. 
Since we invoke LEM only finitely many times, we can use their finite conjunction to prove $\neg\neg$LLPO. 
The double negations of those instances are provable in Type Theory \todo{(include the proof of this?)}.
We can leverage the fact that double negation is a modality, to construct a function that assumes the double negations of the LEM instances and concludes $\neg\neg$LLPO.
\end{proof}

\section{unassigned}

\begin{defn}
The Independence Principle for $X : \mathcal{U}$ and family of propositions $Q: X \rightarrow N \rightarrow \mathtt{Prop}$ is the following statement
$$\mathtt{IP}_{X,Q} : \Big(\prod_{x : X}\| \Sigma_{n : \mathbb{N}} Q(x,n) \| \Big) \rightarrow \| \Sigma_{n : \mathbb{N}}\prod_{x : X} Q(x,n) \|$$
\end{defn}
We know that there exists model of extensional type theory that validates IP. 
We would like a variant of IP to which we can apply Theorem 6.1 of CTCA. 
We can then reach the desired form by omitting the truncation in the antecedent of IP. 
\begin{defn}
We define $\mathtt{IP'}_{X,Q}$ with $X : \mathcal{U}$ and family of propositions $Q: X \rightarrow N \rightarrow \mathtt{Prop}$ 
$$\mathtt{IP}_{X,Q} : \Big(\prod_{x : X} \Sigma_{n : \mathbb{N}} Q(x,n)  \Big) \rightarrow \| \Sigma_{n : \mathbb{N}}\prod_{x : X} Q(x,n) \|$$
\end{defn}
Firstly, $\mathtt{IP}_{X,Q} \Rightarrow \mathtt{IP'}_{X,Q}$. 
Secondly, $\mathtt{IP'}$ is strong enough to prove that $\mathbb{N}$ is $B$-null, where 
$$B : A \rightarrow \mathcal{U}$$
$$S :\equiv \mathbb{N} \rightarrow \mathbf{2}$$
$$A : \Sigma_{s : S}\mathtt({atMost1one}\; s)$$
$$B : \prod_{a : A}\|\mathtt{p_{odd}}\; a.\mathtt{fst} + \mathtt{p_{even}}\; a.\mathtt{fst}\|$$
Specifically, it is enough to make sure that 
$$\prod_{a : A}\prod_{f : B'(a) \rightarrow \mathbb{N}} \mathtt{IP}_{B'(a),Q(a, f)}$$
where $B'(a)$ is the untruncation of $B(a)$ and
$$Q(a, f) : \prod_{b : B'(a)} \prod_{n : \mathbb{N}} \mathcal{U}$$
and is defined as 
$$Q(a, f)\; b\; n \equiv f(b) =_{\mathbb{N}} n$$
